### 2021-11-29 newvariant.tex
am I done?

Inferring the relative strength (i.e., the ratio of reproduction numbers) and relative speed (i.e., the difference between growth rates) of new SARS-CoV-2 variants compared to earlier strains is critical to predicting and controlling the course of the current pandemic.
Here, we challenge a common assumption (constant relative speed) in estimating the transmission advantage of SARS-CoV-2 variants.
We argue that holding relative strength constant is generally more appropriate by comparing how relative strength and speed change over time under two idealized interventions:
a constant-strength intervention like idealized vaccination and social distancing, which reduces transmission rates by a constant amount, and a constant-speed intervention like idealized contact tracing, which isolate infected individuals at a constant rate.
The impact of these interventions on the spread of new variants is evaluated using a generation-interval-based framework in light of recent evidence that the Alpha and Delta variants may have different generation intervals from prior dominant lineages.
Differences in the generation-interval distributions affect not only the effectiveness of these interventions but also the realized transmission advantages of new variants.
Neglecting these differences gives biased estimates of relative strength, which can vary systematically with changes in the incidence of infections caused by the variant.

@@

Here, we use the generation-interval-based framework to compare two measures of relative transmissibility of new variants: the relative strength ($\rho=\Rv/\Rw$) and relative speed ($\delta=\rv-\rw$).
We assess how relative strength and speed depend on underlying epidemiologicaldynamics of prior dominant lineages and argue that assuming a constant relative strength (rather than a constant relative speed) is more appropriate for estimating relative transmissibility of new variants.
We also show how neglecting differences in the generation-interval distributions of new variants can lead to biased estimates of their relative transmissibility and how such biases might be assessed in practice.
Finally, we discuss how information on differences in generation interval distributions might influence priorities for controlling the spread of VoCs.

@@

We explored how the generation-interval distribution shapes the link between relative strength and speed of an invading pathogen variant.
Longer generation intervals generally lead to higher relative strength for a given relative speed (and conversely, lower relative speed for a given relative strength); these relationships are reversed when incidence is decreasing.
Neglecting potential differences in the mean generation intervals between the variant and the wild type can bias estimates of the relative strength from incidence data;
these biases may be assessed by considering whether estimates of relative strength appear to vary systematically with the direction of changes in the incidence of infections caused by the variant.
Finally, differences in generation intervals can also lead to different conclusions about the effectiveness of interventions.
If the variant has longer generation intervals than the wild type, speed-like interventions will be relatively more effective than naive estimates would suggest. 
Conversely, strength-like intervention will be relatively more effective if the variant has shorter generation intervals.

@@

Even though SARS-CoV-2 has been spreading for more than a year, there is still considerable uncertainty about its generation intervals.
Several studies have tried to estimate the generation-interval distribution, with means ranging between 3--6 days and squared coefficients of variation ranging between 0.1--1 \citep{ferretti2020quantifying,Ferretti2020timing,ganyani2020estimating,knight2020estimating}.
However, these estimates are derived from serial intervals (i.e., time between symptom onset of the infector and the infectee; \cite{svensson2007note}), which are subject to dynamical biases \citep{park2021forward} and fail to account for asymptomatic transmission, adding further uncertainty to inferences of speed and strength \citep{park2020time}.
Future studies should prioritize detailed assessment of the generation intervals of SARS-CoV-2 and widespread variants, as well as consider how uncertainty in generational intervals might bias conclusions \citep{doi:10.1098/rsif.2020.0144,ali2020serial,gostic2020practical}.

@@

\begin{thebibliography}{10}

\bibitem{switzerland2021variant}
Althaus CL, Reichmuth M, Hodcroft E, Riou J, Schibler M, Eckerle I, Kaiser L,
  Suter F, Huber M, Trkola A, and others .
\newblock 2021 {Transmission of SARS-CoV-2 variants in Switzerland}.
\newblock \url{https://ispmbern.github.io/covid-19/variants/}.

\bibitem{davies2021estimated}
Davies NG, Abbott S, Barnard RC, Jarvis CI, Kucharski AJ, Munday JD, Pearson
  CAB, Russell TW, Tully DC, Washburne AD, and others .
\newblock 2021 {Estimated transmissibility and impact of SARS-CoV-2 lineage
  B.1.1.7 in England}.
\newblock {\em Science}.

\bibitem{di2021impact}
Di~Domenico L, Pullano G, Sabbatini CE, L{\'e}vy-Bruhl D, and Colizza V.
\newblock 2021 {Impact of January 2021 social distancing measures on SARS-CoV-2
  B. 1.1. 7 circulation in France}.
\newblock {\em medRxiv}.
\newblock \url{https://www.medrxiv.org/content/10.1101/2021.02.14.21251708v1}.

\bibitem{graham2021changes}
Graham MS, Sudre CH, May A, Antonelli M, Murray B, Varsavsky T, Kl{\"a}ser K,
  Canas LS, Molteni E, and others .
\newblock 2021/05/03 2021 {Changes in symptomatology, reinfection, and
  transmissibility associated with the SARS-CoV-2 variant B.1.1.7: an
  ecological study}.
\newblock {\em Lancet Public Health}.
\newblock {\bfseries 6}, e335--e345.

\bibitem{leung2021early}
Leung K, Shum MH, Leung GM, Lam TT, and Wu~JT.
\newblock 2021 {Early transmissibility assessment of the N501Y mutant strains
  of SARS-CoV-2 in the United Kingdom, October to November 2020}.
\newblock {\em Euro Surveill.}
\newblock {\bfseries 26}, 2002106.

\bibitem{volz2021transmission}
Volz E, Mishra S, Chand M, Barrett JC, Johnson R, Geidelberg L, Hinsley WR,
  Laydon DJ, Dabrera G, O'Toole {\'A}, and others .
\newblock 2021 {Assessing transmissibility of SARS-CoV-2 lineage B.1.1.7 in
  England}.
\newblock {\em Nature}.

\bibitem{zhao2021}
Zhao S, Lou J, Cao L, Zheng H, Chong MKC, Chen Z, Chan RWY, Zee BCY, Chan PKS,
  and Wang MH.
\newblock 01 2021 {Quantifying the transmission advantage associated with N501Y
  substitution of SARS-CoV-2 in the UK: an early data-driven analysis}.
\newblock {\em J. Travel Med}.
\newblock taab011.

\bibitem{campbell2021increased}
Campbell F, Archer B, Laurenson-Schafer H, Jinnai Y, Konings F, Batra N, Pavlin
  B, Vandemaele K, Van~Kerkhove MD, Jombart T, and others .
\newblock 2021 Increased transmissibility and global spread of {SARS-CoV-2}
  variants of concern as at {June}, 2021.
\newblock {\em Euro Surveill.}
\newblock {\bfseries 26}, 2100509.

\bibitem{diekmann1990definition}
Diekmann O, Heesterbeek JAP, and Metz JA.
\newblock 1990 On the definition and the computation of the basic reproduction
  ratio $\mathcal{R}_0$ in models for infectious diseases in heterogeneous
  populations.
\newblock {\em J. Math. Biol.}
\newblock {\bfseries 28}, 365--382.

\bibitem{anderson1991infectious}
Anderson RM and May RM.
\newblock 1991 {\em {Infectious Diseases of Humans: Dynamics and Control}}.
\newblock Oxford University Press.

\bibitem{mills2004transmissibility}
Mills CE, Robins JM, and Lipsitch M.
\newblock 2004 Transmissibility of 1918 pandemic influenza.
\newblock {\em Nature}.
\newblock {\bfseries 432}, 904--906.

\bibitem{nishiura2009transmission}
Nishiura H, Castillo-Chavez C, Safan M, and Chowell G.
\newblock 2009 {Transmission potential of the new influenza A (H1N1) virus and
  its age-specificity in Japan}.
\newblock {\em Euro Surveill.}
\newblock {\bfseries 14}, 19227.

\bibitem{ma2014estimating}
Ma~J, Dushoff J, Bolker BM, and Earn DJD.
\newblock 2014 Estimating initial epidemic growth rates.
\newblock {\em Bull. Math. Biol.}
\newblock {\bfseries 76}, 245--260.

\bibitem{wallinga2007generation}
Wallinga J and Lipsitch M.
\newblock 2007 How generation intervals shape the relationship between growth
  rates and reproductive numbers.
\newblock {\em Proc. R. Soc. B}.
\newblock {\bfseries 274}, 599--604.

\bibitem{svensson2007note}
Svensson {\AA}.
\newblock 2007 A note on generation times in epidemic models.
\newblock {\em Math. Biosci.}
\newblock {\bfseries 208}, 300--311.

\bibitem{fine2003interval}
Fine PE.
\newblock 2003 The interval between successive cases of an infectious disease.
\newblock {\em Am. J. Epidemiol.}
\newblock {\bfseries 158}, 1039--1047.

\bibitem{grassly2008mathematical}
Grassly NC and Fraser C.
\newblock 2008 Mathematical models of infectious disease transmission.
\newblock {\em Nat. Rev. Microbiol.}
\newblock {\bfseries 6}, 477--487.

\bibitem{britton2019estimation}
Britton T and Scalia~Tomba G.
\newblock 2019 {Estimation in emerging epidemics: Biases and remedies}.
\newblock {\em J. R. Soc. Interface}.
\newblock {\bfseries 16}, 20180670.

\bibitem{ali2020serial}
Ali ST, Wang L, Lau EH, Xu~XK, Du~Z, Wu~Y, Leung GM, and Cowling BJ.
\newblock 2020 {Serial interval of SARS-CoV-2 was shortened over time by
  nonpharmaceutical interventions}.
\newblock {\em Science}.
\newblock {\bfseries 369}, 1106--1109.

\bibitem{park2021forward}
Park SW, Sun K, Champredon D, Li~M, Bolker BM, Earn DJ, Weitz JS, Grenfell BT,
  and Dushoff J.
\newblock 2021 Forward-looking serial intervals correctly link epidemic growth
  to reproduction numbers.
\newblock {\em Proc. Natl. Acad. Sci. U.S.A.}
\newblock {\bfseries 118}, e2011548118.

\bibitem{lloyd2001realistic}
Lloyd AL.
\newblock 2001 Realistic distributions of infectious periods in epidemic
  models: changing patterns of persistence and dynamics.
\newblock {\em Theor. Popul. Biol.}
\newblock {\bfseries 60}, 59--71.

\bibitem{wearing2005appropriate}
Wearing HJ, Rohani P, and Keeling MJ.
\newblock 2005 Appropriate models for the management of infectious diseases.
\newblock {\em PLoS medicine}.
\newblock {\bfseries 2}, e174.

\bibitem{roberts2007model}
Roberts M and Heesterbeek J.
\newblock 2007 Model-consistent estimation of the basic reproduction number
  from the incidence of an emerging infection.
\newblock {\em J. Math. Biol.}
\newblock {\bfseries 55}, 803.

\bibitem{huber2016quantitative}
Huber JH, Johnston GL, Greenhouse B, Smith DL, and Perkins TA.
\newblock 2016 Quantitative, model-based estimates of variability in the
  generation and serial intervals of \textit{Plasmodium falciparum} malaria.
\newblock {\em Malaria journal}.
\newblock {\bfseries 15}, 1--12.

\bibitem{miller2010epidemics}
Miller JC, Davoudi B, Meza R, Slim AC, and Pourbohloul B.
\newblock 2010 Epidemics with general generation interval distributions.
\newblock {\em J. Theor. Biol.}
\newblock {\bfseries 262}, 107--115.

\bibitem{svensson2015influence}
Svensson {\AA}.
\newblock 2015 The influence of assumptions on generation time distributions in
  epidemic models.
\newblock {\em Math. Biosci.}
\newblock 270, 81--89.

\bibitem{park2019practical}
Park SW, Champredon D, Weitz JS, and Dushoff J.
\newblock 2019 A practical generation-interval-based approach to inferring the
  strength of epidemics from their speed.
\newblock {\em Epidemics}.
\newblock 27, 12--18.

\bibitem{eaton2014proportion}
Eaton JW and Hallett TB.
\newblock 2014 Why the proportion of transmission during early-stage {HIV}
  infection does not predict the long-term impact of treatment on {HIV}
  incidence.
\newblock {\em Proc. Natl. Acad. Sci. U.S.A.}
\newblock {\bfseries 111}, 16202--16207.

\bibitem{powers2014impact}
Powers KA, Kretzschmar ME, Miller WC, and Cohen MS.
\newblock 2014 Impact of early-stage {HIV} transmission on treatment as
  prevention.
\newblock {\em Proc. Natl. Acad. Sci. U.S.A.}
\newblock {\bfseries 111}, 15867--15868.

\bibitem{weitz2015modeling}
Weitz JS and Dushoff J.
\newblock 2015 Modeling post-death transmission of {Ebola}: challenges for
  inference and opportunities for control.
\newblock {\em Sci. Rep.}
\newblock {\bfseries 5}, 1--7.

\bibitem{park2020time}
Park SW, Cornforth DM, Dushoff J, and Weitz JS.
\newblock 2020 The time scale of asymptomatic transmission affects estimates of
  epidemic potential in the {COVID-19} outbreak.
\newblock {\em Epidemics}.
\newblock 31, 100392.

\bibitem{doi:10.1098/rspb.2020.1556}
Dushoff J and Park SW.
\newblock 2021 Speed and strength of an epidemic intervention.
\newblock {\em Proc. R. Soc. B}.
\newblock {\bfseries 288}, 20201556.

\bibitem{kissler2021densely}
Kissler S, Fauver JR, Mack C, Tai CG, Breban MI, Watkins AE, Samant RM,
  Anderson DJ, Ho~DD, Grubaugh ND, and Grad YH.
\newblock 2021 {Densely sampled viral trajectories suggest longer duration of
  acute infection with B.1.1.7 variant relative to non-B.1.1.7 SARS-CoV-2}.
\newblock {\em medRxiv}.
\newblock \url{https://www.medrxiv.org/content/10.1101/2021.02.16.21251535v1}.

\bibitem{li2021viral}
Li~B, Deng A, Li~K, Hu~Y, Li~Z, Xiong Q, Liu Z, Guo Q, Zou L, Zhang H, and
  others .
\newblock 2021 {Viral infection and Transmission in a large well-traced
  outbreak caused by the Delta SARS-CoV-2 variant}.
\newblock {\em medRxiv}.
\newblock \url{https://www.medrxiv.org/content/10.1101/2021.07.07.21260122v1}.

\bibitem{zhang2021transmission}
Zhang M, Xiao J, Deng A, Zhang Y, Zhuang Y, Hu~T, Li~J, Tu~H, Li~B, Zhou Y, and
  others .
\newblock 2021 Transmission dynamics of an outbreak of the {COVID-19} {Delta}
  variant {B.1.617.2}--{Guangdong Province, China, May--June 2021}.
\newblock {\em China CDC Weekly}.
\newblock 3, 584--586.

\bibitem{goldstein2009reconstructing}
Goldstein E, Dushoff J, Ma~J, Plotkin JB, Earn DJ, and Lipsitch M.
\newblock 2009 Reconstructing influenza incidence by deconvolution of daily
  mortality time series.
\newblock {\em Proc. Natl. Acad. Sci. U.S.A.}
\newblock {\bfseries 106}, 21825--21829.

\bibitem{gostic2020practical}
Gostic KM, McGough L, Baskerville EB, Abbott S, Joshi K, Tedijanto C, Kahn R,
  Niehus R, Hay JA, De~Salazar PM, and others .
\newblock 12 2020 Practical considerations for measuring the effective
  reproductive number, $\mathcal{R}_t$.
\newblock {\em PLOS Comp. Biol.}
\newblock {\bfseries 16}, 1--21.

\bibitem{heesterbeek1996concept}
Heesterbeek J and Dietz K.
\newblock 1996 The concept of $\mathcal{R}_0$ in epidemic theory.
\newblock {\em Stat. Neerl.}
\newblock {\bfseries 50}, 89--110.

\bibitem{diekmann2000mathematical}
Diekmann O and Heesterbeek JAP.
\newblock 2000 {\em Mathematical Epidemiology of Infectious Diseases: Model
  Building, Analysis and Interpretation}, volume~5.
\newblock John Wiley \& Sons.

\bibitem{roberts2004modelling}
Roberts M.
\newblock 2004 Modelling strategies for minimizing the impact of an imported
  exotic infection.
\newblock {\em Proc. R. Soc. B}.
\newblock {\bfseries 271}, 2411--2415.

\bibitem{aldis2005integral}
Aldis G and Roberts M.
\newblock 2005 An integral equation model for the control of a smallpox
  outbreak.
\newblock {\em Math. Biosci.}
\newblock {\bfseries 195}, 1--22.

\bibitem{breda2012formulation}
Breda D, Diekmann O, De~Graaf W, Pugliese A, and Vermiglio R.
\newblock 2012 {On the formulation of epidemic models (an appraisal of Kermack
  and McKendrick)}.
\newblock {\em J. Biol. Dyn.}
\newblock {\bfseries 6}, 103--117.

\bibitem{champredon2018equivalence}
Champredon D, Dushoff J, and Earn DJ.
\newblock 2018 Equivalence of the {Erlang}-distributed {SEIR} epidemic model
  and the renewal equation.
\newblock {\em SIAM J. Appl. Math}.
\newblock {\bfseries 78}, 3258--3278.

\bibitem{fraser2007estimating}
Fraser C.
\newblock 2007 Estimating individual and household reproduction numbers in an
  emerging epidemic.
\newblock {\em PloS one}.
\newblock {\bfseries 2}.

\bibitem{champredon2015intrinsic}
Champredon D and Dushoff J.
\newblock 2015 Intrinsic and realized generation intervals in
  infectious-disease transmission.
\newblock {\em Proc. R. Soc. B}.
\newblock {\bfseries 282}, 20152026.

\bibitem{champredon2018two}
Champredon D, Li~M, Bolker BM, and Dushoff J.
\newblock 2018 Two approaches to forecast {Ebola} synthetic epidemics.
\newblock {\em Epidemics}.
\newblock 22, 36--42.

\bibitem{kenah2008generation}
Kenah E, Lipsitch M, and Robins JM.
\newblock 2008 Generation interval contraction and epidemic data analysis.
\newblock {\em Math. Biosci.}
\newblock {\bfseries 213}, 71--79.

\bibitem{nishiura2010time}
Nishiura H.
\newblock 2010 Time variations in the generation time of an infectious disease:
  implications for sampling to appropriately quantify transmission potential.
\newblock {\em Math. Biosci. Eng.}
\newblock {\bfseries 7}, 851.

\bibitem{flaxman2020Rt}
Flaxman S, Mishra S, Gandy A, Unwin HJT, Mellan TA, Coupland H, Whittaker C,
  Zhu H, Berah T, Eaton JW, Monod M, and others .
\newblock 2020 {Estimating the effects of non-pharmaceutical interventions on
  COVID-19 in Europe}.
\newblock {\em Nature}.
\newblock {\bfseries 584}, 257--261.

\bibitem{moore2021vaccination}
Moore S, Hill EM, Tildesley MJ, Dyson L, and Keeling MJ.
\newblock 2021 {Vaccination and non-pharmaceutical interventions for COVID-19:
  a mathematical modelling study}.
\newblock {\em Lancet Infect. Dis.}
\newblock {\bfseries 21}, 793--802.

\bibitem{brauner2021inferring}
Brauner JM, Mindermann S, Sharma M, Johnston D, Salvatier J, Gaven{\v{c}}iak T,
  Stephenson AB, Leech G, Altman G, Mikulik V, and others .
\newblock 2021 {Inferring the effectiveness of government interventions against
  COVID-19}.
\newblock {\em Science}.
\newblock {\bfseries 371}.

\bibitem{doi:10.1098/rsif.2020.0144}
Park SW, Bolker BM, Champredon D, Earn DJD, Li~M, Weitz JS, Grenfell BT, and
  Dushoff J.
\newblock 2020 {Reconciling early-outbreak estimates of the basic reproductive
  number and its uncertainty: framework and applications to the novel
  coronavirus (SARS-CoV-2) outbreak}.
\newblock {\em J. R. Soc. Interface}.
\newblock {\bfseries 17}, 20200144.

\bibitem{ferretti2020quantifying}
Ferretti L, Wymant C, Kendall M, Zhao L, Nurtay A, Abeler-D{\"o}rner L, Parker
  M, Bonsall D, and Fraser C.
\newblock 2020 {Quantifying SARS-CoV-2 transmission suggests epidemic control
  with digital contact tracing}.
\newblock {\em Science}.
\newblock {\bfseries 368}.

\bibitem{leung2017monitoring}
Leung K, Lipsitch M, Yuen KY, and Wu~JT.
\newblock 2017 Monitoring the fitness of antiviral-resistant influenza strains
  during an epidemic: a mathematical modelling study.
\newblock {\em Lancet Infect. Dis.}
\newblock {\bfseries 17}, 339--347.

\bibitem{leung2020empirical}
Leung K, Pei Y, Leung GM, Lam TT, and Wu~JT.
\newblock 2020 {Empirical transmission advantage of the D614G mutant strain of
  SARS-CoV-2}.
\newblock {\em medRxiv}.
\newblock \url{https://www.medrxiv.org/content/10.1101/2020.09.22.20199810v1}.

\bibitem{wallinga2004different}
Wallinga J and Teunis P.
\newblock 2004 Different epidemic curves for severe acute respiratory syndrome
  reveal similar impacts of control measures.
\newblock {\em Am. J. Epidemiol.}
\newblock {\bfseries 160}, 509--516.

\bibitem{cori2013new}
Cori A, Ferguson NM, Fraser C, and Cauchemez S.
\newblock 2013 A new framework and software to estimate time-varying
  reproduction numbers during epidemics.
\newblock {\em Am. J. Epidemiol.}
\newblock {\bfseries 178}, 1505--1512.

\bibitem{abbott2020estimating}
Abbott S, Hellewell J, Thompson RN, Sherratt K, Gibbs HP, Bosse NI, Munday JD,
  Meakin S, Doughty EL, Chun JY, and others .
\newblock 2020 {Estimating the time-varying reproduction number of SARS-CoV-2
  using national and subnational case counts}.
\newblock {\em Wellcome Open Research}.
\newblock {\bfseries 5}, 112.

\bibitem{knight2020estimating}
Knight J and Mishra S.
\newblock 2020 {Estimating effective reproduction number using generation time
  versus serial interval, with application to COVID-19 in the Greater Toronto
  Area, Canada}.
\newblock {\em Infect. Dis. Model.}
\newblock 5, 889--896.

\bibitem{li2021temporal}
Li~Y, Campbell H, Kulkarni D, Harpur A, Nundy M, Wang X, Nair H, for COVID UN,
  and others .
\newblock 2021 {The temporal association of introducing and lifting
  non-pharmaceutical interventions with the time-varying reproduction number
  (R) of SARS-CoV-2: a modelling study across 131 countries}.
\newblock {\em Lancet Infect. Dis.}
\newblock {\bfseries 21}, 193--202.

\bibitem{tsang2020effect}
Tsang TK, Wu~P, Lin Y, Lau EH, Leung GM, and Cowling BJ.
\newblock 2020 {Effect of changing case definitions for COVID-19 on the
  epidemic curve and transmission parameters in mainland China: a modelling
  study}.
\newblock {\em The Lancet Public Health}.
\newblock {\bfseries 5}, e289--e296.

\bibitem{ukinvest}
{Public Health England} .
\newblock 2020 {Investigation of novel SARS-CoV-2 variant: Variant of Concern
  202012/01}.
\newblock
  \url{https://assets.publishing.service.gov.uk/government/uploads/system/uploads/attachment_data/file/959426/Variant_of_Concern_VOC_202012_01_Technical_Briefing_5.pdf}.

\bibitem{ganyani2020estimating}
Ganyani T, Kremer C, Chen D, Torneri A, Faes C, Wallinga J, and Hens N.
\newblock 2020 {Estimating the generation interval for coronavirus disease
  (COVID-19) based on symptom onset data, March 2020}.
\newblock {\em Euro Surveill.}
\newblock {\bfseries 25}, 2000257.

\bibitem{fraser2004factors}
Fraser C, Riley S, Anderson RM, and Ferguson NM.
\newblock 2004 Factors that make an infectious disease outbreak controllable.
\newblock {\em Proc. Natl. Acad. Sci. U.S.A.}
\newblock {\bfseries 101}, 6146--6151.

\bibitem{scarabel2021renewal}
Scarabel F, Pellis L, Ogden NH, and Wu~J.
\newblock 2021 A renewal equation model to assess roles and limitations of
  contact tracing for disease outbreak control.
\newblock {\em R. Soc. Open Sci.}
\newblock {\bfseries 8}, 202091.

\bibitem{unwin2020state}
Unwin HJT, Mishra S, Bradley VC, Gandy A, Mellan TA, Coupland H, Ish-Horowicz
  J, Vollmer MA, Whittaker C, Filippi SL, and others .
\newblock 2020 {State-level tracking of COVID-19 in the United States}.
\newblock {\em Nat. Commun.}
\newblock {\bfseries 11}, 1--9.

\bibitem{brett2020transmission}
Brett TS and Rohani P.
\newblock 2020 Transmission dynamics reveal the impracticality of {COVID-19}
  herd immunity strategies.
\newblock {\em Proc. Natl. Acad. Sci. U.S.A.}
\newblock {\bfseries 117}, 25897--25903.

\bibitem{Ferretti2020timing}
Ferretti L, Ledda A, Wymant C, Zhao L, Ledda V, Abeler-D{\"o}rner L, Kendall M,
  Nurtay A, Cheng HY, Ng~TC, and others .
\newblock 2020 The timing of {COVID-19} transmission.
\newblock {\em medRxiv}.
\newblock \url{https://www.medrxiv.org/content/10.1101/2020.09.04.20188516v2}.

\bibitem{abdool2021new}
Abdool~Karim SS and de~Oliveira T.
\newblock 2021 {New SARS-CoV-2 variants—clinical, public health, and vaccine
  implications}.
\newblock {\em N. Engl. J. Med.}

\bibitem{fontanet2021sars}
Fontanet A, Autran B, Lina B, Kieny MP, Karim SSA, and Sridhar D.
\newblock 2021 {SARS-CoV-2 variants and ending the COVID-19 pandemic}.
\newblock {\em Lancet}.
\newblock {\bfseries 397}, 952--954.

\bibitem{walensky2021sars}
Walensky RP, Walke HT, and Fauci AS.
\newblock 2021 {SARS-CoV-2 variants of concern in the United
  States---Challenges and opportunities}.
\newblock {\em JAMA}.
\newblock {\bfseries 325}, 1037--1038.

\bibitem{hellewell2020feasibility}
Hellewell J, Abbott S, Gimma A, Bosse NI, Jarvis CI, Russell TW, Munday JD,
  Kucharski AJ, Edmunds WJ, Sun F, and others .
\newblock 2020 {Feasibility of controlling COVID-19 outbreaks by isolation of
  cases and contacts}.
\newblock {\em Lancet Glob. Health}.
\newblock {\bfseries 8}, e488--e496.

\end{thebibliography}

>>>>

Inferring the relative strength (i.e., the ratio of reproduction numbers) and relative speed (i.e., the difference between growth rates) of new SARS-CoV-2 variants compared to earlier strains is critical to predicting and controlling the course of the current pandemic.
Here, we challenge a commonly-made assumption (i.e., a constant relative speed) in estimating the transmission advantage of SARS-CoV-2 variants.
We argue that holding relative strength constant is generally more appropriate by comparing how relative strength and speed change over time under two idealized interventions:
a constant-strength intervention like idealized vaccination and social distancing, which reduces transmission rates by a constant amount, and a constant-speed intervention like idealized contact tracing, which isolate infected individuals at a constant rate.
The impact of these interventions on the spread of new variants is evaluated using a generation-interval-based framework in light of recent evidence that the Alpha and Delta variants may have different generation intervals from prior dominant lineages.
Differences in the generation-interval distributions affect not only the effectiveness of these interventions but also the realized transmission advantages of new variants.
Neglecting these differences gives biased estimates of relative strength, which can vary systematically with changes in the incidence of infections caused by the variant.

@@

Here, we use the generation-interval-based framework to compare two measures of transmission advantage of new variants: the relative strength ($\rho=\Rv/\Rw$) and relative speed ($\delta=\rv-\rw$).
We assess how relative strength and speed depend on underlying epidemiological dynamics of prior dominant lineages and argue that assuming a constant relative strength (rather than a constant relative speed) is more appropriate for estimating relative transmissibility of new variants.
We also show how neglecting differences in the generation-interval distributions of new variants can lead to biased estimates of their relative transmissibility and how such biases might be assessed in practice.
Finally, we discuss how information on differences in generation interval distributions might influence priorities for controlling the spread of VoCs.

@@

We explored how the generation-interval distribution shapes the link between relative strength and speed.
While the role of generation-interval distributions in linking epidemic strength and speed has been well-established, using the generation-interval-based framework provides clear expanations for an imporant, but under-appreciated concept: the relative strength and speed of new variants also depend on epidemic dynamics of the prior dominant lineages as well as the type of control measures (i.e., strength- or speed-like) that limit their spread.
For example, under a constant-strength intervention that reduces the transmission rate of the wild type and a variant by the same amount (and therefore, keeping the relative strength constant), the relative speed of the variant is sensitive to the epidemic growth rate of the wild type strain---this challenges the commonly-made assumption that relative speed of new variants remains constant over time.
Differences in the generation-interval distributions further exaggerate this effect.

Therefore, neglecting potential differences in the mean generation intervals between the variant and the wild type can bias estimates of the relative strength from incidence data.
These biases may be assessed by considering whether estimates of relative strength appear to vary systematically with the direction of changes in the incidence of infections caused by the variant.
Differences in generation intervals can also lead to different conclusions about the effectiveness of interventions.
If the variant has longer generation intervals than the wild type, speed-like interventions will be relatively more effective than naive estimates would suggest. 
Conversely, strength-like intervention will be relatively more effective if the variant has shorter generation intervals.

@@

Even though SARS-CoV-2 has been spreading for more than a year, there is still considerable uncertainty about its generation intervals.
Several studies have tried to estimate the generation-interval distribution, with means ranging between 3--6 days and squared coefficients of variation ranging between 0.1--1 \citep{ferretti2020quantifying,Ferretti2020timing,ganyani2020estimating,knight2020estimating}.
However, these estimates are derived from serial intervals (i.e., time between symptom onset of the infector and the infectee; \cite{svensson2007note}), which are subject to dynamical biases \citep{park2021forward} and fail to account for asymptomatic transmission \citep{park2020time}.
Recently, it has been further suggested that the generation intervals of the original SARS-CoV-2 strain may be considerably longer than previously thought \citep{sender2021unmitigated}, adding further uncertainty.
Although a few studies have tried to quantify both the generation- and serial-interval distributions of new SARS-CoV-2 variants, their conclusions on whether transmission time scale of new variants differs are conflicting \citep{pung2021serial, hart2021generation}---we note that these comparisons are inherently difficult especially due to temporal changes in generation- and serial-interval distributions due to dynamical and intervention effects \citep{ali2020serial,park2021forward}.
Future studies should prioritize detailed assessment of the generation intervals of SARS-CoV-2 and widespread variants, as well as consider how uncertainty in generational intervals might bias conclusions \citep{doi:10.1098/rsif.2020.0144,ali2020serial,gostic2020practical}.

@@

\bibliography{newvariant}
